---
title: "Homework_5"
format: html
authors: Evan Whitfield, Ralph Church, Ryan Friedman
---

## Task 1 - Conceptual Questions

Create a markdown list with the following questions:

### Question 1

What is the purpose of the `lapply()` function? What is the equivalent `purrr` function?

**Answer:** The equivalent `purrr` function is `map()`.

### Question 2

Suppose we have a list called my_list. Each element of the list is a numeric data frame (all columns are numeric). We want use `lapply()` to run the code `cor(numeric_matrix, method = "kendall")` on each element of the list. Write code to do this below! (I’m really trying to ask you how you specify `method = "kendall"` when calling `lapply()`)

**Answer:**

### Question 3

What are two advantages of using `purrr` functions instead of the BaseR `apply` family?

**Answer:**

### Question 4

What is a side-effect function?

**Answer:**

### Question 5

Why can you name a variable `sd` in a function and not cause any issues with the `sd` function?

**Answer:**


## Task 2 - Writing R Functions

### Step 1

**Task:** Write a basic function (call it `getRMSE()`) that takes in a vector of responses and a vector of predictions and outputs the RMSE (Root Mean Square Error).

```{r}
getRMSE <- function(obs, pre, ...) {
  if (length(obs) != length(pre)){
    stop("Dimension mis-match")
  }
  MSE <- mean((obs - pre)^2, ...)
  RMSE <- sqrt(MSE)
  RMSE
}
```

```{r}
getRMSE(c(1,2,3,NA), c(2,2,2,4), na.rm = TRUE)
```

### Step 2

**Task:** Run the following code to create some response values and predictions.

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

```

```{r}
getRMSE(resp, pred)
```

```{r}
resp[1] <- NA
pred[4] <- NA

resp #checking to see 1st index is NA
```

```{r}
#Should return NA do to NA in the vectors.
getRMSE(resp, pred)
```

```{r}
#Should return a value since NAs have been removed.
getRMSE(resp, pred, na.rm = TRUE)
```


### Step 3

**Task:** Write a function called `getMAE()` that follows the specifications of the `getRMSE()` function.

```{r}
getMAE <- function(obs, pre, ...) {
  if (length(obs) != length(pre)){
    stop("Dimension mis-match")
  }
  MAE = mean(abs(obs - pre), ...)
  MAE
}
```

### Step 4

**Task:** Run the following code to create some response values and predictions.

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

```{r}
getMAE(resp, pred)
```
```{r}
resp[1] <- NA
pred[4] <- NA

#Testing to see if NA shows in resp
resp
```

```{r}
#Should return NA
getMAE(resp, pred)
```

```{r}
#Should return a value
getMAE(resp, pred, na.rm = TRUE)
```

### Step 5

**Task:** Let’s create a `wrapper` function that can be used to get either or both metrics returned with a single function call. Do not rewrite your above two functions, call them inside the `wrapper` function (we would call the `getRMSE()` and `getMAE()` functions helper functions). When returning your values, give them appropriate names.

```{r}
#change function name
wrapper <- function(.x, .y) {
  if (FALSE) {
    stop("Incorrect data passed as an argument.")
  }
  5
}
```


### Step 6

**Task:** Run the following code to create some response values and predictions.

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

