---
title: "Homework_5"
format: html
authors: Evan Whitfield, Ralph Church, Ryan Friedman
---

## Task 1 - Conceptual Questions

Create a markdown list with the following questions:

### Question 1

What is the purpose of the `lapply()` function? What is the equivalent `purrr` function?

**Answer:** Use `lapply` to apply functions to each element in a list. The equivalent `purrr` function is `map()`.

### Question 2

Suppose we have a list called my_list. Each element of the list is a numeric data frame (all columns are numeric). We want use `lapply()` to run the code `cor(numeric_matrix, method = "kendall")` on each element of the list. Write code to do this below! (I’m really trying to ask you how you specify `method = "kendall"` when calling `lapply()`)

**Answer:** lapply(my_list, cor, method = "kendall")

### Question 3

What are two advantages of using `purrr` functions instead of the BaseR `apply` family?

**Answer:** 
- The `purrr` functions tell you exactly what type of output you'll get, whereas the `apply` family can sometimes be difficult to predict the type of output. 

- The `purrr` functions also work with the `tidyverse` which makes for cleaner and more readable syntax.

### Question 4

What is a side-effect function?

**Answer:** A side-effect function is a function that does something rather than `return` a value. It may print to the console, plot a graph, create a file, etc.  

### Question 5

Why can you name a variable `sd` in a function and not cause any issues with the `sd` function?

**Answer:** The variable inside of the function and the sd function would live in different environments. R looks for the named object/function in the first environment, and then moves on the 2nd environment if not found in the first environment. Values inside of functions exist in the first level, then the .GlobalEnv, and then the different packages loaded into R, until package:base.


## Task 2 - Writing R Functions

### Step 1

**Task:** Write a basic function (call it `getRMSE()`) that takes in a vector of responses and a vector of predictions and outputs the RMSE (Root Mean Square Error).

```{r}
getRMSE <- function(obs, pre, ...) {
  if (length(obs) != length(pre)){
    stop("Dimension mis-match")
  }
  MSE <- mean((obs - pre)^2, ...)
  RMSE <- sqrt(MSE)
  RMSE
}
```

Our function checks to see if the vector of observed and predicted values is the same length. Then, it calculates and returns the RMSE. NA values can be removed if necessary.


### Step 2

**Task:** Run the following code to create some response values and predictions.

In order to test our `getRMSE()` function, we create a vector of observed values called `resp`, and a vector of prediction values called `pred`. We will follow this approach for each function.

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

Without any missing values, our function should return a value for RMSE.
```{r}
getRMSE(resp, pred)
```
The RMSE is found to be 0.9581677.

```{r}
resp[1] <- NA
pred[4] <- NA

#Should return NA do to NA in the vectors.
getRMSE(resp, pred)
```
With NA values present in either the observed vector or predicted vector (we test NAs in both), the function returns NA.

```{r}
#Should return a value since NAs have been removed.
getRMSE(resp, pred, na.rm = TRUE)
```
Removing the NA values from the vectors, the function is able to calcualte the RMSE once again.

### Step 3

**Task:** Write a function called `getMAE()` that follows the specifications of the `getRMSE()` function.

```{r}
getMAE <- function(obs, pre, ...) {
  if (length(obs) != length(pre)){
    stop("Dimension mis-match")
  }
  MAE = mean(abs(obs - pre), ...)
  MAE
}
```

Now, we can calculate the MAE just like we did for the RMSE.

### Step 4

**Task:** Run the following code to create some response values and predictions.

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

```{r}
getMAE(resp, pred)
```
Without any NA values in either vector, the MAE is 0.8155776.

```{r}
resp[1] <- NA
pred[4] <- NA

#Should return NA
getMAE(resp, pred)
```
Once again, with NA values present in either vector, the function returns NA.

```{r}
#Should return a value
getMAE(resp, pred, na.rm = TRUE)
```
If we remove any NA values, we can calculate the new MAE.

### Step 5

**Task:** Let’s create a `wrapper` function that can be used to get either or both metrics returned with a single function call. Do not rewrite your above two functions, call them inside the `wrapper` function (we would call the `getRMSE()` and `getMAE()` functions helper functions). When returning your values, give them appropriate names.

We will need to use the helper function below to check to see if the correct type of data is being passed into our wrapper function.
```{r}
is.numeric_vector <- function(.x) {
  if (is.numeric(.x) && is.vector(.x) && is.atomic(.x)) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}
```

We create our `getMETRICS()` function to able to return any specified metrics. As of now, it only works for RMSE and MAE, and defaults to finding both.
```{r}
library(purrr)
getMETRICS <- function(obs, 
                       pre, 
                       metrics = c("RMSE","MAE"),
                       ...) {
  
  #Scale for new functions/metrics to be added later
  metric_funs <- list(
    RMSE = getRMSE,
    MAE  = getMAE
  )
  
  #Checking to see that the inputs are correct for our function.
  if (!(("RMSE" %in% metrics) || ("MAE" %in% metrics))) {
    stop("Either RMSE or MAE must be passed in metrics.")
  }
  if (any(!metrics %in% names(metric_funs))) {
    stop("Unknown metric passed in metrics.")
  }
  if (!(is.numeric_vector(obs) && is.numeric_vector(pre))) {
    stop("Incorrect data passed as an argument. Please use numeric vectors.")
  }

  #Map the list of metrics we want to calculate.
  result <- map_dbl(metrics, 
                \(metr) metric_funs[[metr]](obs, pre, ...)
  )

  #Add names to reference the correct metric.
  names(result) <- metrics
  result
}
```


### Step 6

**Task:** Run the following code to create some response values and predictions.

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

```{r}
getMETRICS(resp,pred)
```
We can see that the function works, and gives both desired metrics.
```{r}
getMETRICS(resp,pred)["RMSE"]
```
We can call a single metric by name if we would like.

```{r}
getMETRICS(resp,pred, metrics = "RMSE")
```
Alternatively, we can pass the necessary metric as an argument.

```{r}
getMETRICS(resp,pred, metrics = "MAE")
```

```{r}
resp[1] <- NA
pred[4] <- NA
getMETRICS(resp,pred)
```
The function returns NA for metrics if NA values are present in the vectors.

```{r}
getMETRICS(resp,pred,na.rm=TRUE)
```

If we remove the NA values from the vectors, then once again, our function will give us the RMSE and MAE metrics.
