---
title: "RC_HW5"
format: html
---

## Task 1: Conceptual Questions

1.  What is the purpose of the `lapply()` function? What is the equivalent `purrr` function?

-   `lapply()` applies a function to each element of a list and returns the results as a list. The equivalent `purrr` function is `map()`. The `purrr` package also includes variations like `map_chr()`, `map_dbl()`, and `map_df()` when a specific output type is needed.

2.  Suppose we have a list called `my_list`. Each element of the list is a numeric data frame. We want to use `lapply()` to run `cor(numeric_matrix, method = "kendall")` on each element of the list. Write code to do this below.

-   `lapply(my_list, cor, method = "kendall")`

3.  What are two advantages of using `purrr` functions instead of the Base R apply family?

-   `purrr` functions have more consistent syntax and work well with tidyverse workflows.
-   `purrr` has type-specific functions, such as `map_dbl()` and `map_chr()`, which make the expected output clearer.

4.  What is a side-effect function?

-   A side-effect function performs an action without mainly returning a transformed object. Examples include printing output, creating plots, writing files, or reading data.

5.  Why can you name a variable `sd` in a function and not cause any issues with the `sd()` function?

-   Functions have their own local environments. A variable named `sd` inside a function only exists within that function and does not overwrite the base R `sd()` function globally.

## Task 2: Writing R Functions

1.  Write a basic function called `getRMSE()` that takes in a vector of responses and a vector of predictions and outputs the RMSE. If a value is missing, allow additional arguments to be passed to `mean()` using ellipses.

```{r}
library(purrr)
library(tidyr)
library(dplyr)


getRMSE <- function(resp, pred, ...) {
  if (length(resp) != length(pred)) {
    stop("resp and pred must have the same length")
  }

  sqrt(mean((resp - pred)^2, ...))
}
```

2.  Run the following code to create some response values and predictions.

```{r}
set.seed(10)

n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

Test the RMSE function using this data.

```{r}
getRMSE(resp, pred)
```

Replace two values with missing values and test again.

```{r}
resp[2] <- NA_real_
pred[5] <- NA_real_

getRMSE(resp, pred)
getRMSE(resp, pred, na.rm = TRUE)
```

3.  Write a function called `getMAE()` that follows the specifications of the `getRMSE()` function.

```{r}
getMAE <- function(resp, pred, ...) {
  if (length(resp) != length(pred)) {
    stop("resp and pred must have the same length")
  }

  mean(abs(resp - pred), ...)
}
```

4.  Run the following code to create some response values and predictions.

```{r}
set.seed(10)

n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

Test the MAE function using this data.

```{r}
getMAE(resp, pred)
```

Replace two values with missing values and test again.

```{r}
resp[3] <- NA_real_
pred[4] <- NA_real_

getMAE(resp, pred)
getMAE(resp, pred, na.rm = TRUE)
```

5.  Create a function to get both metrics returned in a single function call.

The function should:

-   check that two numeric atomic vectors have been passed
-   return both metrics by default
-   include names for the returned metrics

```{r}
getBOTH <- function(resp, pred, metrics = c("RMSE", "MAE"), ...) {
  if (!is.atomic(resp) || !is.atomic(pred)) {
    stop("resp and pred must be atomic vectors")
  }

  if (!is.numeric(resp) || !is.numeric(pred)) {
    stop("resp and pred must be numeric")
  }

  if (length(resp) != length(pred)) {
    stop("resp and pred must have the same length")
  }

  valid_metrics <- c("RMSE", "MAE")

  if (!all(metrics %in% valid_metrics)) {
    stop('metrics must contain only "RMSE" and/or "MAE"')
  }

  metric_functions <- list(
    RMSE = getRMSE,
    MAE = getMAE
  )

  results <- map_dbl(metrics, \(metric) {
    metric_functions[[metric]](resp, pred, ...)
  })

  names(results) <- metrics
  results
}
```

6.  Run the following code to create some response values and predictions.

```{r}
set.seed(10)

n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

Test the function using this data. Call it for each metric individually, and then with both.

```{r}
getBOTH(resp, pred, metrics = "RMSE")
getBOTH(resp, pred, metrics = "MAE")
getBOTH(resp, pred, metrics = c("RMSE", "MAE"))
```

Replace two values with missing values, then test again using `na.rm = TRUE`.

```{r}
resp[1] <- NA_real_
pred[6] <- NA_real_

getBOTH(resp, pred, metrics = "RMSE", na.rm = TRUE)
getBOTH(resp, pred, metrics = "MAE", na.rm = TRUE)
getBOTH(resp, pred, metrics = c("RMSE", "MAE"), na.rm = TRUE)
```

Test the function by passing incorrect data.

```{r}
getBOTH(as_tibble(resp), pred)
```

```{r}
getBOTH(as.character(resp), pred)
```

```{r}
getBOTH(list(resp), pred)
```

