---
title: "HW5RF"
format: html
---

## Task 1: Conceptual Questions

1.  What is the purpose of the `lapply()` function? What is the equivalent `purrr` function?

- `lapply()` applies a function to all elements of a list, and returns a list result. This is more convenient than looping through the list elements to apply the function. Within `purrr`, `map()` can be used to accomplish the same thing. `map()` has many variations for returning specific data types and/or accepting multiple list inputs.

2.  Suppose we have a list called my_list. Each element of the list is a numeric data frame (all columns are numeric). We want use lapply() to run the code cor(numeric_matrix, method = "kendall") on each element of the list. Write code to do this below! (I’m really trying to ask you how you specify method = "kendall" when calling lapply())

- `lapply(my_list, cor, method = "kendall")`

3.  What are two advantages of using purrr functions instead of the BaseR apply family?

- `purrr` functions have many options for what inputs to accept and what format of output to return (i.e. `map_chr()` will always return a character vector)

- `purrr` functions are compatible with the tidyverse and have consistent syntax

4.  What is a side-effect function?

- A side-effect function is a function that *does not* modify and return the argument (as opposed to a transformation function which does). Examples are plots or reading data.

5.  Why can you name a variable sd in a function and not cause any issues with the sd function?

- Functions create their own temporary environments for executing code. Variables defined within the function will not overwrite external variables/functions of the same name.

## Task 2: Writing R Functions

First load potentially useful packages.

```{r}
library(purrr)
library(tidyr)
library(dplyr)
```

1.  Write a basic function (call it getRMSE()) that takes in a vector of responses and a vector of predictions and outputs the RMSE. If a value is missing for the vector of responses (i.e. an NA is present), allow for additional arguments to the mean() function (elipses) that removes the NA values in the computation.

```{r}
getRMSE <- function(resp, pred, ...){
  if(length(resp) != length(pred)) stop("vectors must be equal in length") # confirm vectors are equal in length
  diffsq <- (resp-pred)^2 # squared difference
  return(sqrt(mean(diffsq, ...))) # ... allows for na.rm = TRUE as an argument for mean
}
```

2.  Run the following code to create some response values and predictions.

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

- Test the RMSE function using this data.

```{r}
getRMSE(resp, pred)
```

- Replace two response values with missing values (`NA_real_`) and test again.

```{r}
resp[2] <- NA_real_
pred[5] <- NA_real_
getRMSE(resp, pred) # confirm that NA values will cause a NA result
getRMSE(resp, pred, na.rm = TRUE) # confirm that na.rm = TRUE argument successfully passes into the mean() and the RMSE is returned
```

3.  Write a function called getMAE() (mean absolute deviation) that follows the specifications of the getRMSE() function.

```{r}
getMAE <- function(resp, pred, ...){
  if(length(resp) != length(pred)) stop("vectors must be equal in length") # confirm vectors are equal in length
  diffabs <- abs(resp-pred) # absolute difference
  return(mean(diffabs, ...)) # ... allows for na.rm = TRUE as an argument for mean
}
```

4.  Run the following code to create some response values and predictions.

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

- Test the MAE function using this data.

```{r}
getMAE(resp, pred)
```

- Replace two response values with missing values (`NA_real_`) and test again.

```{r}
resp[3] <- NA_real_
pred[4] <- NA_real_
getMAE(resp, pred) # confirm that NA values will cause a NA result
getMAE(resp, pred, na.rm = TRUE) # confirm that na.rm = TRUE argument successfully passes into the mean() and the RMSE is returned
```

5.  Create a function to get both metrics returned in a single function call.

- The function should check that two numeric atomic vectors have been passed

- The function should return both metrics by default and include names

```{r}
getBOTH <- function(resp, pred, metrics = c("RMSE", "MAE"), ...){
  # error checking the two vector inputs
  if(!is.vector(resp)|!is.vector(pred)) stop("resp and pred must be vectors")
  if(!is.atomic(resp)|!is.atomic(pred)) stop("resp and pred must be atomic (i.e. not lists)")
  if(!is.numeric(resp)|!is.numeric(pred)) stop("resp and pred must be numeric")
  # error checking the metric request
  if(!(all(metrics %in% c("RMSE", "MAE")))) stop("metrics must be a string vector of \"RMSE\" and/or \"MAE\"")
  # make a lookup table of desired functions
  metFunctions <- list(
    RMSE = getRMSE,
    MAE = getMAE
  )
  # build results list based on metrics requested
  results <- map(metrics, \(x) metFunctions[[x]](resp, pred, ...))
  # assign names
  names(results) <- metrics
  return(results)
}
```

6.  Run the following code to create some response values and predictions.

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

- Test the function using this data. Call it for each metric individually, and then with both.

```{r}
getBOTH(resp, pred, metrics = "RMSE")
getBOTH(resp, pred, metrics = "MAE")
getBOTH(resp, pred, metrics = c("RMSE", "MAE"))
```

- Replace two of the response values with missing values, then test again. Use na.rm = TRUE

```{r}
resp[1] <- NA_real_
pred[6] <- NA_real_
getBOTH(resp, pred, metrics = "RMSE", na.rm = TRUE)
getBOTH(resp, pred, metrics = "MAE", na.rm = TRUE)
getBOTH(resp, pred, metrics = c("RMSE", "MAE"), na.rm = TRUE)
```

- Test function by passing incorrect data

```{r}
getBOTH(as_tibble(resp), pred)
```

```{r}
getBOTH(as.character(resp), pred)
```

```{r}
getBOTH(list(resp), pred)
```
